---
title: "Exercise Quality"
author: "Lineesh Thamaran"
date: "Monday, October 24, 2016"
output: html_document
---

##Introducton:
People regularly collect large amount of data about their personal health activitties and quantify how much of a particular activity they do. This project tries to quantify the quality of their activities.

##Analysis:
The manner in which the activities are performed is in the "classes" variable. Our goal is to predict the "classe" variable for the pml_testing data set.

##Preprocess:
Download the training and test data into dataTrain and dataTest csv files.

```{r}
library(caret)
trainUrl <- 'https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv'
testUrl <- 'https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv'
download.file(url = trainUrl, destfile = 'data_train.csv')
download.file(url = testUrl, destfile = 'data_test.csv')
```

Preliminary inspection shows that some columns contain data that is not useful. Load the data in memory, and remove data that is not useful. 
```{r}
dataTrain <- read.csv(file = 'data_train.csv', na.strings = c('NA','#DIV/0!',''))
dataTest <- read.csv(file = 'data_test.csv', na.strings = c('NA','#DIV/0!',''))
```

Explore the data.
```{r}
summary(dataTrain)
levels(dataTrain$classe)
```

Exploratory data analysis shows that the first 7 columns are not useful for prediction model. Convert the rest of the columns (except "classe") to numeric.
```{r}
for(i in c(8:ncol(dataTrain)-1)) {
  dataTrain[,i] = as.numeric(as.character(dataTrain[,i]))
  dataTest[,i] = as.numeric(as.character(dataTest[,i]))
}
```

Create a usedColumns vector with the columns that we are planning to use. i.e remove columns with NA in it. Also remove the first 7 columns.
```{r}
usedColumns <- colnames(dataTrain)
usedColumns <- colnames(dataTrain[colSums(is.na(dataTrain)) == 0])
usedColumns <- usedColumns[-c(1:7)]
```

Plot the features against "classe". The plot shows similar distribution for all features. 
```{r}
featurePlot(dataTrain[,usedColumns[c(1:10)]], dataTrain[,160], "strip")
```

Partition the data into training and test sets. 80% of data is used for training and 20% for testing.
```{r}
set.seed(32343)
index_train <- createDataPartition(y=dataTrain$classe, p=0.80, list=FALSE)
p_data_train <- dataTrain[index_train,usedColumns]
p_data_val <- dataTrain[-index_train,usedColumns]
```

##Training
Train the model using random forest and create the confusion matrix:
```{r}
mod_rf <- train(classe ~ ., data = p_data_train, method = 'rf', trControl = trainControl(method = "cv", number = 4, allowParallel = TRUE, verboseIter = TRUE))
pred_rf <- predict(mod_rf, p_data_val)
cm_rf <- confusionMatrix(pred_rf, p_data_val$classe)
cm_rf
```

Accuracy of our model is 0.9949. So the sample error is 0.0051 (i.e 1 - accuracy). Since our goal is to predict the outcome for 20 cases, 99% accuracy is sufficent. 

##Prediction:
Predict the classe for the test data using the model.
```{r}
last_column <- length(colnames(dataTest[]))
colnames(dataTest)[last_column] <- 'classe'
quiz_rf <- predict(mod_rf,dataTest[,usedColumns])
quiz_rf
```





